Final Ratings for Genome Biology

### 2018-10-26 manuscript_figures/manuscript_ratings.Rmd
Create supplementary plot for ratings

```{r load-workspace, message = FALSE}
## Load packages and source benchmark FDR
library(tidyverse)
library(cowplot)
library(ggthemes)
library(grid)
##library(SummarizedBenchmark)
devtools::load_all("../../SummarizedBenchmark")

@@

## Robustness

Next, we define metrics for the robustness metric, which quantifies the amount of consistency,
or robustness, of each method across simulations, yeast experiments, and case studies. For
this metric, we only look at modern methods, i.e. exclude Bonferroni correction, BH, and
Storey's q-value. We do this, as we quantify "robustness" as the relative gain of each method
over the average of BH and q-value in FDR and TPR (and proportion of rejections for case studies).

```{r}
## robustness across simulation settings
robustness_sims <- simres %>%
    dplyr::filter(key == "TPR" | key == "FDR", alpha == alpha.thresh,
                  !grepl("null", setting), sdist != "chisq",
                  Method != "bonf") %>%
    dplyr::select(Method, key, mean.info, setting, sdist, inform, pi0, signal, ntests) %>%
    tidyr::spread(Method, mean.info) %>%
    tidyr::gather(Method, mean.info, -key, -setting, -sdist, -inform,
                  -pi0, -signal, -ntests, -qvalue, -bh) %>%
    dplyr::group_by(Method, key) %>%
    dplyr::summarize(meanGain = mean(mean.info / (bh + qvalue) * 2),
                     sdGainBH = sd(mean.info / bh),
                     sdGainQ = sd(mean.info / qvalue),
                     sdGain = sd(mean.info / (bh + qvalue) * 2),
                     cvGain = sdGain / meanGain) %>%
    dplyr::ungroup()

## robustness across yeast experiments
robustness_yeast <- yeast %>%
    dplyr::filter(key == "TPR" | key == "FDR", !grepl("null", setting_de)) %>%
    dplyr::select(Method, mean, key, setting_de, setting_cov, setting_mode, setting_pi0) %>%
    tidyr::spread(Method, mean) %>%
    tidyr::gather(Method, mean, -key, -setting_de, -setting_cov, -setting_pi0,
                  -setting_mode, -qvalue, -bh) %>%
    dplyr::group_by(Method, key) %>%
    dplyr::summarize(meanGain = mean(mean / (bh + qvalue) * 2),
                     sdGainBH = sd(mean / bh),
                     sdGainQ = sd(mean / qvalue),
                     sdGain = sd(mean / (bh + qvalue) * 2),
                     cvGain = sdGain / meanGain) %>%
    dplyr::ungroup()

## robustness across case studies
robustness_data <- datres %>%
    dplyr::mutate(prej = nrej / ntotal) %>%
    dplyr::select(-rank, -nrej, -ntotal, -nNA) %>%
    tidyr::spread(Method, prej) %>%
    tidyr::gather(Method, prej, -casestudy, -dataset, -qvalue, -bh) %>%
    dplyr::group_by(Method) %>%
    dplyr::summarize(meanGain = mean(prej / (bh + qvalue) * 2, na.rm = TRUE),
                     sdGainBH = sd(prej / bh, na.rm = TRUE),
                     sdGainQ = sd(prej / qvalue, na.rm = TRUE),
                     sdGain = sd(prej / (bh + qvalue) * 2, na.rm = TRUE),
                     cvGain = sdGain / meanGain) %>%
    dplyr::ungroup() %>%
    dplyr::mutate(key = "PropRej")

## combine
robustness <- dplyr::full_join(dplyr::select(robustness_sims, Method, key, cvGain),
                               dplyr::select(robustness_yeast, Method, key, cvGain),
                               by = c("key", "Method"),
                               suffix = c(".sims", ""))
robustness <- dplyr::full_join(robustness,
                               dplyr::select(robustness_data, Method, key, cvGain),
                               by = c("key", "Method"),
                               suffix = c(".yeast", ".data"))
dplyr::arrange(robustness, key, cvGain.sims + cvGain.yeast, cvGain.data)
```

@@

```{r, fig.width = 6, fig.height = 6}
bind_rows(yeast = robustness_yeast, sims = robustness_sims,
          casestudies = robustness_data, .id = "analysis") %>%
    dplyr::mutate(Method = reorder(Method, sdGain)) %>%
    ggplot(aes(x = Method, y = sdGain, color = Method, group = analysis)) +
    geom_line(aes(linetype = analysis), color = 'black', alpha = 1/4) +
    geom_point() + 
    scale_color_manual(values = col, guide = FALSE) +
    scale_x_discrete(NULL) + 
    scale_y_continuous("SD of percent relative to BH/qvalue") + 
    expand_limits(y = 1) +
    facet_grid(key ~ ., scales = "free") + 
    theme_classic() +
    theme(axis.title = element_text(face = "bold", size = 10),
          plot.title = element_text(face = "bold", size = 14)) +
    ggtitle("Variation across simulations, yeast experiments,\ncase studies")

bind_rows(yeast = robustness_yeast, sims = robustness_sims,
          casestudies = robustness_data, .id = "analysis") %>%
    dplyr::mutate(Method = reorder(Method, cvGain)) %>%
    ggplot(aes(x = Method, y = cvGain, color = Method, group = analysis)) +
    geom_line(aes(linetype = analysis), color = 'black', alpha = 1/4) +
    geom_point() + 
    scale_color_manual(values = col, guide = FALSE) +
    scale_x_discrete(NULL) + 
    scale_y_continuous("CV of percent relative to BH/qvalue") + 
    expand_limits(y = 0) +
    facet_grid(key ~ ., scales = "free") + 
    theme_classic() +
    theme(axis.title = element_text(face = "bold", size = 10),
          plot.title = element_text(face = "bold", size = 14)) +
    ggtitle("Variation across simulations, yeast experiments,\ncase studies")

bind_rows(yeast = robustness_yeast, sims = robustness_sims,
          casestudies = robustness_data, .id = "analysis") %>%
    dplyr::mutate(Method = reorder(Method, cvGain)) %>%
    ggplot(aes(x = Method, y = cvGain, color = Method, group = key)) +
    geom_line(aes(linetype = key), color = 'black', alpha = 1/4) +
    geom_point() + 
    scale_color_manual(values = col, guide = FALSE) +
    scale_x_discrete(NULL) + 
    scale_y_continuous("CV of percent relative to BH/qvalue") +
    scale_linetype("Metric") + 
    expand_limits(y = 0) +
    facet_grid(analysis ~ ., scales = "free_y") + 
    theme_classic() +
    theme(axis.title = element_text(face = "bold", size = 10),
          plot.title = element_text(face = "bold", size = 14)) +
    ggtitle("Variation across simulations, yeast experiments,\ncase studies")

bind_rows(yeast = robustness_yeast, sims = robustness_sims,
          casestudies = robustness_data, .id = "analysis") %>%
    dplyr::mutate(Method = reorder(Method, sdGain)) %>%
    ggplot(aes(x = Method, y = sdGain, color = Method, group = key)) +
    geom_line(aes(linetype = key), color = 'black', alpha = 1/4) +
    geom_point() + 
    scale_color_manual(values = col, guide = FALSE) +
    scale_x_discrete(NULL) + 
    scale_y_continuous("SD of percent relative to BH/qvalue") +
    scale_linetype("Metric") + 
    expand_limits(y = 0) +
    facet_grid(analysis ~ ., scales = "free_y") + 
    theme_classic() +
    theme(axis.title = element_text(face = "bold", size = 10),
          plot.title = element_text(face = "bold", size = 14)) +
    ggtitle("Variation across simulations, yeast experiments,\ncase studies") 
```

@@

>>>>

```{r load-workspace, message = FALSE}
## Load packages and source benchmark FDR
library(tidyverse)
library(cowplot)
library(ggthemes)
library(grid)
library(SummarizedBenchmark)

@@

## Consistency

Next, we define metrics for the consistency metric, which quantifies the amount of consistency,
of each method across simulations and case studies. While the yeast experiments
could also be included as part of the consistency calculation, it is excluded, primarily because
all settings are variations on differential expression in RNA-seq, and therefore contains far
less variable settings as compared to the simulations and case studies.

For this metric, we only look at modern methods, i.e. exclude Bonferroni correction, BH, and
Storey's q-value. We do this, as we quantify "consistency" as the relative gain of each method
over the average of BH and q-value in FDR and TPR (and proportion of rejections for case studies).

We exclude both `ashq` and `fdrreg-t` from the analysis of the case studies here because they
were applied in substantially fewer case studies (only 4/26). The gain over BH/q-value can vary
substantially based on the informativeness of the covariate and therefore, if a method was applied
to far fewer case studies, it may have an arbitrarily smaller or larger amount of variability across
case studies simply due to the composition of the subset relative to the composition of the
entire collection of case studies.

```{r}
## consistency across simulation settings
consistency_sims <- simres %>%
    dplyr::filter(key == "TPR" | key == "FDR", alpha == alpha.thresh,
                  !grepl("null", setting), sdist != "chisq",
                  Method != "bonf") %>%
    dplyr::select(Method, key, mean.info, setting, sdist, inform, pi0, signal, ntests) %>%
    tidyr::spread(Method, mean.info) %>%
    tidyr::gather(Method, mean.info, -key, -setting, -sdist, -inform,
                  -pi0, -signal, -ntests, -qvalue, -bh) %>%
    dplyr::group_by(Method, key) %>%
    dplyr::summarize(meanGain = mean(mean.info / (bh + qvalue) * 2),
                     meanRGain = mean(log2(mean.info / (bh + qvalue) * 2)),
                     sdGainBH = sd(mean.info / bh),
                     sdGainQ = sd(mean.info / qvalue),
                     sdGain = sd(mean.info / (bh + qvalue) * 2),
                     sdRGain = sd(log2(.001 + mean.info / (bh + qvalue) * 2)),
                     cvGain = sdGain / meanGain) %>%
    dplyr::ungroup()

## consistency across case studies
consistency_data <- datres %>%
    dplyr::filter(Method != "ashq", Method != "fdrreg-t", !is.na(nrej)) %>%
    dplyr::mutate(prej = (nrej+1) / ntotal) %>%
    dplyr::select(-rank, -nrej, -ntotal, -nNA) %>%
    tidyr::spread(Method, prej) %>%
    tidyr::gather(Method, prej, -casestudy, -dataset, -qvalue, -bh) %>%
    dplyr::group_by(Method) %>%
    dplyr::summarize(meanGain = mean(prej / (bh + qvalue) * 2, na.rm = TRUE),
                     meanRGain = mean(log2(prej / (bh + qvalue) * 2), na.rm = TRUE),
                     sdGainBH = sd(prej / bh, na.rm = TRUE),
                     sdGainQ = sd(prej / qvalue, na.rm = TRUE),
                     sdGain = sd(prej / (bh + qvalue) * 2, na.rm = TRUE),
                     sdRGain = sd(log2(prej / (bh + qvalue) * 2), na.rm = TRUE),
                     cvGain = sdGain / meanGain) %>%
    dplyr::ungroup() %>%
    dplyr::mutate(key = "%reject")
```

@@

```{r, fig.width = 6, fig.height = 6}
bind_rows(Sims = consistency_sims, CS = consistency_data, .id = "analysis") %>%
    dplyr::mutate(Method = reorder(Method, sdGain),
                  key = factor(paste0(key, " (", analysis, ")")),
                  key = factor(key, levels = levels(key)[c(2, 3, 1)])) %>%
    ggplot(aes(x = Method, y = sdGain, color = Method, group = key)) +
    geom_line(aes(linetype = key), color = 'black', alpha = 1/2) +
    geom_point() + 
    scale_color_manual(values = col, guide = FALSE) +
    scale_x_discrete(NULL) + 
    scale_y_continuous("SD of % relative to BH/qvalue") + 
    expand_limits(y = 1) +
    theme_classic() +
    theme(axis.title = element_text(face = "bold", size = 10),
          plot.title = element_text(face = "bold", size = 14)) +
    ggtitle("Variation across simulations and case studies")

bind_rows(Sims = consistency_sims, CS = consistency_data, .id = "analysis") %>%
    dplyr::mutate(Method = reorder(Method, cvGain),
                  key = factor(paste0(key, " (", analysis, ")")),
                  Metric = factor(key, levels = levels(key)[c(2, 3, 1)])) %>%
    ggplot(aes(x = Method, y = cvGain, color = Method, group = Metric)) +
    geom_line(aes(linetype = Metric), color = 'black', alpha = 1/2) +
    geom_point() + 
    scale_color_manual(values = col, guide = FALSE) +
    scale_x_discrete(NULL) + 
    scale_y_continuous("CV of % relative to BH/qvalue") + 
    expand_limits(y = 1) +
    theme_classic() +
    theme(axis.title = element_text(face = "bold", size = 10),
          plot.title = element_text(face = "bold", size = 14)) +
    ggtitle("Variation across simulations and case studies")

bind_rows(Sims = consistency_sims, CS = consistency_data, .id = "analysis") %>%
    dplyr::mutate(Method = reorder(Method, sdRGain),
                  key = factor(paste0(key, " (", analysis, ")")),
                  key = factor(key, levels = levels(key)[c(2, 3, 1)])) %>% 
    ggplot(aes(x = Method, y = sdRGain, color = Method, group = key)) +
    geom_line(aes(linetype = key), color = 'black', alpha = 1/2) +
    geom_point() + 
    scale_color_manual(values = col, guide = FALSE) +
    scale_x_discrete(NULL) + 
    scale_y_continuous("SD of log-ratio to BH/qvalue") + 
    expand_limits(y = 1) +
    theme_classic() +
    theme(axis.title = element_text(face = "bold", size = 10),
          plot.title = element_text(face = "bold", size = 14)) +
    ggtitle("Variation across simulations and case studies")
```

Additionally, we look at the proportion of times a method performed better than BH and Storey's
q-value.

```{r}
## consistency (definition 2) across simulation settings
consistency2_sims <- simres %>%
    dplyr::filter(key == "TPR", alpha == alpha.thresh,
                  !grepl("null", setting), sdist != "chisq",
                  Method != "bonf") %>%
    dplyr::select(Method, key, mean.info, setting, sdist, inform, pi0, signal, ntests) %>%
    tidyr::spread(Method, mean.info) %>%
    tidyr::gather(Method, mean.info, -key, -setting, -sdist, -inform,
                  -pi0, -signal, -ntests, -qvalue, -bh) %>%
    dplyr::group_by(Method, key) %>%
    dplyr::summarize(propBoth = mean(mean.info < pmax(bh, qvalue)),
                     propBoth95 = mean(mean.info < 0.95 * pmax(bh, qvalue)),
                     propEither = mean(mean.info < pmin(bh, qvalue)),
                     propEither95 = mean(mean.info < 0.95 * pmin(bh, qvalue))) %>%
    dplyr::ungroup()

## consistency (definition 2) across case studies
consistency2_data <- datres %>%
    dplyr::filter(Method != "ashq", Method != "fdrreg-t", !is.na(nrej)) %>%
    dplyr::select(-rank, -ntotal, -nNA) %>%
    tidyr::spread(Method, nrej) %>%
    tidyr::gather(Method, nrej, -casestudy, -dataset, -qvalue, -bh) %>%
    dplyr::group_by(Method) %>%
    dplyr::summarize(propBoth = mean(nrej < pmax(bh, qvalue), na.rm = TRUE),
                     propBoth95 = mean(nrej < 0.95 * pmax(bh, qvalue), na.rm = TRUE),
                     propEither = mean(nrej < pmin(bh, qvalue), na.rm = TRUE),
                     propEither95 = mean(nrej < 0.95 * pmin(bh, qvalue), na.rm = TRUE)) %>%
    dplyr::ungroup() %>%
    dplyr::mutate(key = "%reject")
```

We also plot these results for the lowerbound case of out-performing either classical approach (BH or Storey's q-value).

```{r, fig.width = 6, fig.height = 6}
bind_rows(Sims = consistency2_sims, CS = consistency2_data, .id = "analysis") %>%
    dplyr::mutate(Method = reorder(Method, propEither95)) %>%
    ggplot(aes(x = Method, y = propEither95, color = Method, group = analysis)) +
    geom_line(aes(linetype = analysis), color = 'black', alpha = 1/2) +
    geom_point() + 
    scale_color_manual(values = col, guide = FALSE) +
    scale_x_discrete(NULL) + 
    scale_y_continuous("Percent of settings w/ TPR or #rejections\n <95% of BH/q-value", labels = scales::percent) + 
    theme_classic() +
    theme(axis.title = element_text(face = "bold", size = 10),
          plot.title = element_text(face = "bold", size = 14)) +
    ggtitle("Consistency of gains over BH or q-value\nacross simulations and case studies")
```

@@

# Figures

Figures for the manuscript are also generated.

```{r, fig.width = 10, fig.height = 8}
p1a <- bind_rows(Yeast = fdrcontrol_yeast, Sims = fdrcontrol_sims, .id = "Analysis") %>%
    dplyr::mutate(Method = reorder(Method, pExceeds),
                  Analysis = factor(Analysis, levels = c("Yeast", "Sims"))) %>%
    ggplot(aes(x = Method, y = pExceeds, color = Method, group = Analysis)) +
    geom_line(aes(linetype = Analysis), color = 'black', alpha = 1/2) +
    geom_point(size = 3) + 
    scale_color_manual(values = col, guide = FALSE) +
    scale_x_discrete(NULL) + 
    scale_y_continuous("% settings exceeding nominal FDR", labels = scales::percent) + 
    theme_classic() +
    theme(axis.title = element_text(face = "bold", size = 10),
          plot.title = element_text(face = "bold", size = 14)) +
    ggtitle("FDR Control")

p1b <- bind_rows(Yeast = power_yeast, Sims = power_sims, .id = "Analysis") %>%
    dplyr::mutate(Method = reorder(Method, meanRank),
                  Analysis = factor(Analysis, levels = c("Yeast", "Sims"))) %>%
    ggplot(aes(x = Method, y = meanRank, color = Method, group = Analysis)) +
    geom_line(aes(linetype = Analysis), color = 'black', alpha = 1/2) +
    geom_point(size = 3) + 
    scale_color_manual(values = col, guide = FALSE) +
    scale_x_discrete(NULL) + 
    scale_y_continuous("Mean TPR rank", breaks = 0:10) + 
    expand_limits(y = 1) +
    theme_classic() +
    theme(axis.title = element_text(face = "bold", size = 10),
          plot.title = element_text(face = "bold", size = 14)) +
    ggtitle("Power")

p1c <- bind_rows(Sims = consistency2_sims, CS = consistency2_data, .id = "Analysis") %>%
    dplyr::mutate(Method = reorder(Method, propEither95),
                  Analysis = factor(paste0(Analysis, " (", key, ")")),
                  Aanlysis = factor(Analysis, levels = levels(Analysis)[c(2, 3, 1)])) %>%
    ggplot(aes(x = Method, y = propEither95, color = Method, group = Analysis)) +
    geom_line(aes(linetype = Analysis), color = 'black', alpha = 1/2) +
    geom_point(size = 3) + 
    scale_color_manual(values = col, guide = FALSE) +
    scale_x_discrete(NULL) + 
    scale_y_continuous("% settings w/ metric\n < 95% of BH and q-value", labels = scales::percent) + 
    theme_classic() +
    theme(axis.title = element_text(face = "bold", size = 10),
          plot.title = element_text(face = "bold", size = 14)) +
    ggtitle("Consistency (I)")

p1d <- bind_rows(Sims = consistency_sims, CS = consistency_data, .id = "Analysis") %>%
    dplyr::mutate(Method = reorder(Method, sdRGain),
                  Analysis = factor(paste0(Analysis, " (", key, ")"))) %>%
    ggplot(aes(x = Method, y = sdRGain, color = Method, group = Analysis)) +
    geom_line(aes(linetype = Analysis), color = 'black', alpha = 1/2) +
    geom_point(size = 3) + 
    scale_color_manual(values = col, guide = FALSE) +
    scale_x_discrete(NULL) + 
    scale_y_continuous("SD of log-ratio relative to\nmean of BH and q-value") + 
    expand_limits(y = 1) +
    theme_classic() +
    theme(axis.title = element_text(face = "bold", size = 10),
          plot.title = element_text(face = "bold", size = 14)) +
    ggtitle("Consistency (II)")

p1title <- ggdraw() +
    draw_label("Summary Recommendation Metrics",
               fontface = 'bold')

p1ab <- plot_grid(p1a + guides(linetype = FALSE),
                  p1b + guides(linetype = FALSE),
                  labels = LETTERS[1:2], nrow = 1)
p1ab <- plot_grid(p1ab, get_legend(p1a), nrow = 1, rel_widths = c(1, .2))

p1cd <- plot_grid(p1c + guides(linetype = FALSE),
                  p1d + guides(linetype = FALSE),
                  labels = LETTERS[3:4], nrow = 1)
p1cd <- plot_grid(p1cd, get_legend(p1d), nrow = 1, rel_widths = c(1, .2))

FigS <- plot_grid(p1title, p1ab, p1cd, ncol = 1, rel_heights = c(.15, 1, 1))
FigS

ggsave(file.path(outdir, "SFigureMetrics.pdf"), width=10, height=8)
```
